---
title: "week11"
author: "Chang Dong"
date: "2023-04-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(tidymodels, tidyverse,titanic,ggplot2,discrim)
```
# Linear Discriminant Analysis (LDA)

You are going to fit LDA on the Titanic dataset to see how it compares to the logistic regression you fit in Week 8: Example — Titanic and Yardstick. We used the following code:

```{r}
library( titanic ) 
titanic <- as_tibble( titanic::titanic_train ) 
titanic <- titanic %>% mutate(Survived = factor( Survived ), 
                              Pclass = factor( Pclass ), 
                              Sex = factor( Sex )) %>% 
  dplyr::select(Survived, Pclass, Age, Sex) %>% 
drop_na()
```

What you need to do:

1. Fit LDA on the Titanic dataset using TidyModels. Remember, you are looking to classify whether people survived based on their age, sex, and class.

```{r}
titanic_lda <- discrim_linear( mode = "classification" ) %>% 
  set_engine( "MASS" ) %>% 
  fit( Survived ~ Age + Sex + Pclass, data = titanic ) 
titanic_lda
```

2. Obtain class predictions for the titanic dataset.

```{r}
titanic_preds <- predict( titanic_lda, new_data = titanic ) %>% 
  bind_cols( titanic %>% 
dplyr::select( Survived ) )
```

3. Get the confusion matrix for this model.
```{r}
titanic_preds %>% 
  conf_mat( Survived, .pred_class )
```

4. What is the sensitivity of this model? What is the specificity?
```{r}
tibble( Sensitivity = 361 / (361 + 63), 
        Specificity = 204 / (204 + 86) )
```

5. Obtain probability predictions for this model.
```{r}
titanic_preds <- titanic_preds %>% 
  bind_cols(predict( titanic_lda, 
                     new_data = titanic,
                     type = "prob" )
) 
titanic_preds
```

6. Plot the ROC curve for this model.
```{r}
titanic_preds %>% roc_curve( truth = Survived,
                             estimate = .pred_0 ) %>% 
  autoplot()
```

7. Obtain the AUC for this ROC curve.
```{r}
titanic_preds %>% 
  roc_auc(truth = Survived, estimate = .pred_0)
```

8. Based on this value, would you prefer the logistic regression model or the LDA model to model this data?

#### The AUC for logistic regression is 0.852 > 0.850, so you would only just prefer the logistic regression. It is also much easier to interpret logistic regression, so this makes it even more preferred.

9. Create a 10-fold cross-validation set for this data (strata by Survived).
```{r}
titanic_cv <- vfold_cv( titanic, v = 10, strata = Survived ) 
titanic_cv
```

10. Using the cross-validation sets, do the following:

a. Create a model specification for LDA.

b. Fit LDA on the cross-validation sets.

c. Obtain the estimates of the AUC and accuracy, as well as their standard errors.
```{r}
# a) 
lda_spec <- discrim_linear( mode = "classification" ) %>% 
  set_engine( "MASS" )

# b) 
titanic_lda_resamples <- fit_resamples(
  object = lda_spec, preprocessor = recipe(Survived ~ . , data = titanic), 
  resamples = titanic_cv) 

# c) 
titanic_lda_resamples %>% 
  collect_metrics()
```